---
title: "SSSIHL-CAS: Insurance Fraud Classifier for Health Insurance, a tool for fraud
  detection"
output:
  pdf_document:
    toc: true
  html_notebook: null
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    collapsed: true
    smooth_scroll: true
---

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: yellow;
    border-color: black;
    color: black;
}
</style>


```{r setup, include=FALSE}
# Suppress warnings, messages, prompts, and comments in code output for cleaner presentation
knitr::opts_chunk$set(warning = F,message = F,prompt = F,comment = F)

```


# Project Summary

> Objective:

* This is a tool for insurance fraud detection and prevention using machine learning models. The tool is intended to augment any organizations current Fraud Detection Solution, giving ability to keep pace with the rapidly changing fraud   environment, while minimizing costs and increasing profitability.

> Utility:

* Red flagging claims for possible fraud and further investigation, thereby channelizing the investigation efforts to the   most important/prioritized suspected cases 
* Reducing losses and increasing profitability for insurers / reinsurers 
* Reducing premiums for customer's insurance cover 
* Providing a competitive advantage for the insurers / reinsurers 


# Cleaning the dataset

## Loading required libraries

```{r}
# Load all required libraries for data manipulation, visualization, and machine learning
library(readxl)        # For reading Excel files
library(janitor)       # For data cleaning and hygiene
library(dplyr)         # For data manipulation
library(lubridate)     # For date/time operations
library(caret)         # For machine learning workflows
library(rpart)         # For decision trees
library(rpart.plot)    # For decision tree visualization
library(xgboost)       # For gradient boosting
library(tidyverse)     # Core tidyverse packages
library(randomForest)  # For random forest models
library(pROC)          # For ROC curve analysis
library(e1071)         # For statistical functions including SVM
library(mlbench)       # For machine learning benchmark data
library(naivebayes)    # For Naive Bayes classifier
library(ggplot2)       # For data visualization
library(psych)         # For psychological research tools
library(gbm)           # For generalized boosted regression models
library(caTools)       # For ROC curve calculations
library(glmnet)        # For Lasso/Ridge regression
library(bruceR)        # For data analysis and visualization
library(kableExtra)    # For enhanced table formatting
library(smotefamily)   # For SMOTE oversampling
library(imbalance)     # For handling imbalanced datasets
library(ROSE)          # For ROSE oversampling technique
library(PRROC)         # For precision-recall curves
library(ROCR)          # For ROC performance evaluation

```


# Preparing Input data without triggers

## Loading data and Reducing dimensionality of character columns

```{r}
# Load pre-cleaned input dataset
clean_data1 = readRDS("./data/raw/cleaned_input_data.rds")

## Considering top 20 unique values from certain columns with large unique values 

# medical service provider id - aggregate counts to identify top providers
mspid <- 
clean_data1 %>% 
  group_by(medical_service_provider_id) %>% 
  count() %>% 
  arrange(desc(n))

# Calculate coverage percentage of top 20 providers
sum(mspid[1:20,2])/sum(mspid[,2])

# removing features that do not contribute in predicting fraud
# Focus on top 20 providers and group others, remove ID columns and date fields
data_wo_triggers <-
  clean_data1 %>% 
  mutate(medical_service_provider_id = 
           ifelse(medical_service_provider_id %in% mspid[1:20,1]$medical_service_provider_id,
                  medical_service_provider_id, 
                  "Others" )) %>% 
  dplyr::select(-c(insured_id, primary_diagnosis_code,treatment_start_date,treatment_end_date,policy_commencement_date,policy_termination_date,claim_reported_date,claim_count_pa))



```


## Feature scaling

```{r}
# Scale/normalize numerical features to comparable ranges
k = c("approved_allowed_amount","no_of_days_stayed","birth_date")
data_wo_triggers[k] = scaler(data_wo_triggers[k])

# Save processed dataset for future use
write_rds(data_wo_triggers,"./data/interim/data_wo_triggers.rds",compress = "bz2",compression = 9)

# Display preview of processed data in formatted table
kable(head(data_wo_triggers) , caption = "Data_wo_triggers")%>%
  kable_styling(bootstrap_options = "striped", "hover", "condensed", "responsive", full_width = F, position = "center") %>% 
 row_spec(0, bold = T, color = "black", background = "yellow") %>% 
 scroll_box( height = "900px",fixed_thead = TRUE)

```



## Performing OneHot Encoding to Non-numeric columns

```{r}
# Convert categorical variables to binary dummy variables for machine learning compatibility
dmy <- dummyVars(" ~ .", data = data_wo_triggers, fullRank = TRUE)
data_wo_triggers <- data.frame(predict(dmy, newdata = data_wo_triggers))

```


## checking correlations of cloumns

```{r}
# Calculate correlation matrix to identify multicollinearity issues
corr = cor(data_wo_triggers)

# Simplify correlation matrix by zeroing upper triangle and diagonal
corr[upper.tri(corr)] = 0
diag(corr) = 0
corr = as.data.frame(corr)
corr = corr %>% tibble::rownames_to_column("colnames_1")
piv = pivot_longer(corr,cols = -colnames_1,  names_to = "colnames", values_to = "correlation")
piv = 
  piv %>% 
  mutate(correlation_1 = abs(correlation)) %>% 
  arrange(desc(correlation_1)) %>% 
  filter(correlation_1 !=0)
#selecting features to drop that are highly correlated (threshold: 0.7)
drop = as.vector(subset(piv, correlation_1 > 0.7 , select = colnames))
data_wo_triggers = data_wo_triggers[,!(names(data_wo_triggers) %in% drop$colnames)]
```


### Lasso regression to remove non-influential features in predicting fraud
```{r}

#perform k-fold cross-validation to find optimal lambda value for Lasso regularization
#Lasso (L1) helps select most important features by shrinking coefficients to zero
cv_model <- cv.glmnet(data.matrix(data_wo_triggers %>% dplyr::select(-c("fraud"))), data_wo_triggers$fraud, alpha=1)


#find optimal lambda value that minimizes test MSE
best_lambda = cv_model$lambda.min

#produce plot of test MSE by lambda value for visualization
plot(cv_model)

#find coefficients of best model and select only non-zero coefficients
best_model<-glmnet(data.matrix(data_wo_triggers %>% dplyr::select(-c("fraud"))), data_wo_triggers$fraud, alpha = 1, lambda = best_lambda)
k = as.matrix(coef(best_model))
k = as.data.frame(k[-1,])
k = k %>% rownames_to_column("feature")
names(k)[names(k) == 'k[-1, ]'] <- 'coeff'
# Keep only features with non-zero coefficients
k = k[k$coeff!=0,]
# Sort by absolute coefficient values to find most important features
k = k[order(abs(k$coeff),decreasing = T),]
# Limit to top 30 features if more were selected
if(nrow(k)>30) k = k[1:30,]
# Subset dataset to only selected important features
data_wo_triggers = data_wo_triggers[,c(k$feature,"fraud")]
```


## Checking for class data Imbalance

```{r}
# Examine the distribution of fraud classes to identify imbalance issues
summary(data_wo_triggers$fraud)
barplot(prop.table(table(data_wo_triggers$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")


```

There is evidence to suggest the existence class data imbalance

## SPlitting Train and Test set

```{r}

# randomly split data in r with 70-30 train-test ratio
# Set seed for reproducibility
set.seed(123)
TrainingIndex = sample(seq_len(nrow(data_wo_triggers)),size = 0.7*nrow(data_wo_triggers))
data_wo_triggers_train <- data_wo_triggers[TrainingIndex,] # Training Set
data_wo_triggers_test <- data_wo_triggers[-TrainingIndex,] # Test Set
```


## Adjusting data for imbalance

### ADASYN

```{r}
# Apply ADASYN (Adaptive Synthetic Sampling) to balance minority class
# ADASYN focuses on generating samples for hard-to-learn minority class examples
newData <- ADAS(X = data_wo_triggers_train,data_wo_triggers_train$fraud,K = 5)

adasyn_wo_triggers_train <- newData$data %>% dplyr::select(-c("class"))


# Save balanced training set for model comparison
readr::write_rds(adasyn_wo_triggers_train,"./data/interim/adasyn_wo_triggers_train.rds", compress = "bz2", compression = 9)
adasyn_wo_triggers_train$fraud = as.factor(adasyn_wo_triggers_train$fraud)
# Display class distribution after balancing
print(adasyn_wo_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```



```{r}
# Apply ADASYN to balance test set (for consistent evaluation)
newData <- ADAS(X = data_wo_triggers_test,data_wo_triggers_test$fraud,K = 5)

adasyn_wo_triggers_test <- newData$data %>% dplyr::select(-c("class"))

# Save balanced test set
readr::write_rds(adasyn_wo_triggers_test,"./data/interim/adasyn_wo_triggers_test.rds", compress = "bz2", compression = 9)
adasyn_wo_triggers_test$fraud = as.factor(adasyn_wo_triggers_test$fraud)
# Display class distribution after balancing
print(adasyn_wo_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### SMOTE

```{r}
# Calculate class imbalance ratio to determine oversampling needed
nrow(data_wo_triggers_train)/sum(data_wo_triggers_train$fraud)

# Apply SMOTE (Synthetic Minority Oversampling Technique) to balance classes
# SMOTE creates synthetic samples by interpolating between minority class neighbors
smote_wo_triggers_train <- SMOTE(data_wo_triggers_train,data_wo_triggers_train$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

# Save balanced training set
readr::write_rds(smote_wo_triggers_train,"./data/interim/smote_wo_triggers_train.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_wo_triggers_train$fraud = as.factor(smote_wo_triggers_train$fraud)
# Verify class balance after SMOTE
print(smote_wo_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```



```{r}
# Calculate imbalance ratio for test set
nrow(data_wo_triggers_test)/sum(data_wo_triggers_test$fraud)

# Apply SMOTE to balance test set for consistent evaluation
smote_wo_triggers_test <- SMOTE(data_wo_triggers_test,data_wo_triggers_test$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

readr::write_rds(smote_wo_triggers_test,"./data/interim/smote_wo_triggers_test.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_wo_triggers_test$fraud = as.factor(smote_wo_triggers_test$fraud)
# Verify class balance after SMOTE
print(smote_wo_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```



### MWMOTE

```{r}

# Apply MWMOTE (Majority Weighted Minority Oversampling Technique)
# MWMOTE identifies hard-to-learn minority samples and generates synthetic samples weighted by difficulty
newData <- mwmote(dataset = data_wo_triggers_train,numInstances = nrow(data_wo_triggers_train)-2*sum(data_wo_triggers_train$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

# Combine original data with synthetic samples
mwmote_wo_triggers_train<-rbind(data_wo_triggers_train,newData)

# Save balanced training set
readr::write_rds(mwmote_wo_triggers_train,"./data/interim/mwmote_wo_triggers_train.rds", compress = "bz2", compression = 9)
mwmote_wo_triggers_train$fraud = as.factor(mwmote_wo_triggers_train$fraud)
# Display class distribution after MWMOTE balancing
print(mwmote_wo_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```


```{r}
# Apply MWMOTE to balance test set for consistent evaluation
newData <- mwmote(dataset = data_wo_triggers_test,numInstances = nrow(data_wo_triggers_test)-2*sum(data_wo_triggers_test$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

mwmote_wo_triggers_test<-rbind(data_wo_triggers_test,newData)

readr::write_rds(mwmote_wo_triggers_test,"./data/interim/mwmote_wo_triggers_test.rds", compress = "bz2", compression = 9)
mwmote_wo_triggers_test$fraud = as.factor(mwmote_wo_triggers_test$fraud)
# Display class distribution after MWMOTE balancing
print(mwmote_wo_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### ROSE

```{r}
# Apply ROSE (Random Over-Sampling Examples) for class balancing
# ROSE generates synthetic data in the feature space around minority class observations
#chose seed 100 for reproducibility
#The N variable is the total number of samples that will be generated
#N should essentially be sum of both fraud and non fraud classes
rose_wo_triggers_train <- ovun.sample(fraud~. , data = data_wo_triggers_train, method = "over",N=(nrow(data_wo_triggers_train)-sum(data_wo_triggers_train$fraud))*2,seed = 100)$data

# Save balanced training set
readr::write_rds(rose_wo_triggers_train,"./data/interim/rose_wo_triggers_train.rds", compress = "bz2", compression = 9)
rose_wo_triggers_train$fraud = as.factor(rose_wo_triggers_train$fraud)
# Display and visualize class distribution after ROSE balancing
table(rose_wo_triggers_train$fraud)
barplot(prop.table(table(rose_wo_triggers_train$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```



```{r}
# Apply ROSE to balance test set for consistent evaluation
#chose seed 100 for reproducibility
#The N variable is the total number that will be generated
#N should essentially be sum of both fraud and non fraud classes
rose_wo_triggers_test <- ovun.sample(fraud~. , data = data_wo_triggers_test, method = "over",N=(nrow(data_wo_triggers_test)-sum(data_wo_triggers_test$fraud))*2,seed = 100)$data

readr::write_rds(rose_wo_triggers_test,"./data/interim/rose_wo_triggers_test.rds", compress = "bz2", compression = 9)
rose_wo_triggers_test$fraud = as.factor(rose_wo_triggers_test$fraud)
# Display and visualize class distribution after ROSE balancing
table(rose_wo_triggers_test$fraud)
barplot(prop.table(table(rose_wo_triggers_test$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```


# Preparing input dataset with triggers


## Loading trigger functions

```{r}
# Load custom trigger functions for fraud detection heuristics
source("./src/features/triggers for package.R")

```


## claim_amount_trigger_map

```{r}
# Apply trigger to flag claims exceeding expected package amounts by procedure code
trigger_data <- read_excel("./data/external/trigger_table.xlsx")

# Prepare trigger mapping data
data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, package_amount_trigger)

# Map claim amounts against expected triggers to flag anomalies
data_w_triggers <- claim_amount_trigger_map(claims_file = clean_data1,claim_paid_field = "approved_allowed_amount",triggers_file = data,procedure_code = "primary_procedure_code",procedure_amount_field = "package_amount_trigger")
```


## hosp_days_trigger_map

```{r}
# Apply trigger to flag excessive hospital stay durations by procedure code
data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, admission_days_trigger)
data_w_triggers <- hosp_days_trigger_map(claims_file = data_w_triggers,no_of_days_stayed = "no_of_days_stayed",triggers_file = data,procedure_code = "primary_procedure_code",admission_days_trigger = "admission_days_trigger")
```


## age_trigger_map

```{r}
# Apply trigger to flag claims for procedures not appropriate for patient's age group

data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(lower_age_limit, upper_age_limit, primary_procedure_code)

data$primary_procedure_code <- as.character(data$primary_procedure_code)

data_w_triggers <- age_trigger_map(claims_file = data_w_triggers,age_column_name = "birth_date",triggers_file = data,lower_age_limit = "lower_age_limit", upper_age_limit = "upper_age_limit",procedure_code = "primary_procedure_code")

```

## gender_trigger_map

```{r}
# Apply trigger to flag claims for gender-specific procedures claimed by wrong gender
data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, gender_trigger)

data_w_triggers <- gender_trigger_map(claims_file = data_w_triggers,gender_column_name = "gender_code",triggers_file = data,procedure_code = "primary_procedure_code",gender_trigger = "gender_trigger")
```


## claim_count_trigger_map

```{r}
# Apply trigger to flag excessive claim frequency per annum by procedure code

data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, claim_count_trigger)

data_w_triggers <- claim_count_trigger_map(claim_count = data_w_triggers,claim_count_pa = "claim_count_pa",claim_count_trigger_file = data,procedure_code_field = "primary_procedure_code",claim_count_trigger = "claim_count_trigger")
```


## close_prox_trigger_map

```{r}
# Apply trigger to flag claims filed too soon after policy commencement

data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, close_prox_days)

data_w_triggers <-  close_prox_trigger_map(claims_file = data_w_triggers,treatment_start_date = "treatment_start_date",policy_commencement_date = "policy_commencement_date",triggers_file = data,procedure_code = "primary_procedure_code",close_prox_days = "close_prox_days")
```


## treatment_date_validity_trigger_map

```{r}
# Apply trigger to flag treatments outside policy coverage dates

data_w_triggers <- treatment_date_validity_trigger_map(claims_file = data_w_triggers,treatment_start_date_field = "treatment_start_date",policy_commencement_date_field = "policy_commencement_date",policy_termination_date_field = "policy_termination_date")
```


## claim_reported_delay_trigger_map

```{r}
# Apply trigger to flag claims reported with significant delay (threshold: 15 days)

data_w_triggers <- claim_reported_delay_trigger_map(claims_file = data_w_triggers,treatment_end_date_field = "treatment_end_date",claim_reported_date_field = "claim_reported_date",claim_delay_days = 15)
```


## hospital_empanelled_trigger_map

```{r}
# Apply trigger to flag treatments at non-empanelled hospitals
data <- 
  read_excel("./data/external/empanelled_hospitals_list_masked.xlsx") %>% 
  janitor::clean_names() %>% 
  dplyr::select(hosp_id)

data_w_triggers <- hospital_empanelled_trigger_map(claims_file = data_w_triggers,hospital_id_field = "medical_service_provider_id",empanelled_hospitals_list = data,empanelled_hospital_id = "hosp_id")
```



## hospital_distance_trigger_map

```{r}
# Apply trigger to flag unusual travel distances to hospitals (threshold: 100 km)
data <- 
  read_excel("./data/external/district_distances_masked.xlsx") %>% 
  tidyr::pivot_longer(cols=-"location_1",
                      names_to = "location_2",
                      values_to = "distance")

data_w_triggers <- hospital_distance_trigger_map(claims_file = data_w_triggers,residence_location_field = "residence_location",hospital_location_field = "hospital_location",hospital_distance_file = data,residence_location_map = "location_1",hospital_location_map = "location_2",hospital_distance_field = "distance", distance_threshold = 100)
```


```{r}
# Generate comprehensive summary statistics for dataset with all triggers applied
skimr::skim(data_w_triggers)
```


# Preparing Input data with triggers

## Reducing dimensionality of character columns

```{r}
## Considering top 20 unique values from certain columns with large unique values 

# medical service provider id - aggregate to identify top providers in dataset with triggers
mspid <- 
data_w_triggers %>% 
  group_by(medical_service_provider_id) %>% 
  count() %>% 
  arrange(desc(n))

# Calculate what percentage of data is covered by top 20 providers
sum(mspid[1:20,2])/sum(mspid[,2])

# Reduce dimensionality by grouping less frequent providers together
data_w_triggers <-
  data_w_triggers %>% 
  mutate(medical_service_provider_id = 
           ifelse(medical_service_provider_id %in% mspid[1:20,1]$medical_service_provider_id,
                  medical_service_provider_id, 
                  "Others" )) %>% 
  dplyr::select(-c(insured_id, primary_diagnosis_code,treatment_start_date,treatment_end_date,policy_commencement_date,policy_termination_date,claim_reported_date))


```




## Performing OneHot Encoding to Non-numeric columns

```{r}
# Convert categorical features to binary dummy variables for dataset with triggers
dmy <- dummyVars(" ~ .", data = data_w_triggers, fullRank = TRUE)
data_w_triggers <- data.frame(predict(dmy, newdata = data_w_triggers))
```


## Removing features with SD 0 and removing trigger information features
```{r}
# Remove auxiliary trigger columns, keeping only the derived flag indicators
drop1 = c("package_amount_trigger","admission_days_trigger","lower_age_limit","upper_age_limit","age_flag","gender_triggerFemale","claim_count_trigger","close_prox_days","hospital_empanelled_flag","treatment_date_validity_flag")
data_w_triggers = data_w_triggers %>% select(-c(drop1))
```


## Feature scaling

```{r}
# Scale numerical features including trigger-derived features for proper model training
k = c("approved_allowed_amount","no_of_days_stayed","birth_date","claim_count_pa","claim_duration_days","distance")
data_w_triggers[k] = scaler(data_w_triggers[k])

# Display preview of processed data with triggers
kable(head(data_w_triggers), caption = "data_w_triggers")%>%
  kable_styling(bootstrap_options = "striped", "hover", "condensed", "responsive", full_width = F, position = "center") %>% 
 row_spec(0, bold = T, color = "black", background = "yellow") %>% 
 scroll_box( height = "900px",fixed_thead = TRUE)

# Save processed dataset with triggers
write_rds(data_w_triggers,"./data/interim/data_w_triggers.rds",compress = "bz2",compression = 9)
```


## Checking for class data Imbalance

```{r}
# Check fraud class distribution in dataset with triggers
summary(data_w_triggers$fraud)
barplot(prop.table(table(data_w_triggers$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")


```


## Splitting train and test data
```{r}
# Set seed for reproducibility and align feature sets between datasets
set.seed(123)
# Ensure same features plus trigger-derived features are used
data_w_triggers = data_w_triggers[,c(colnames(data_wo_triggers),"claim_amount_flag","hosp_days_flag","gender_flag","claim_count_flag","close_prox_flag","distance","hospital_distance_flag","claim_duration_days","claim_reported_delay_flag")]

# randomly split data in r with 70-30 train-test ratio
TrainingIndex = sample(seq_len(nrow(data_w_triggers)),size = 0.7*nrow(data_w_triggers))
data_w_triggers_train <- data_w_triggers[TrainingIndex,] # Training Set
data_w_triggers_test <- data_w_triggers[-TrainingIndex,] # Test Set
```



## Adjusting data for imbalance

### ADASYN

```{r}
# Apply ADASYN to dataset WITH trigger features for training set
newData <- ADAS(X = data_w_triggers_train,data_w_triggers_train$fraud,K = 5)

adasyn_w_triggers_train <- newData$data %>% dplyr::select(-c("class"))

# Save balanced training set with triggers
readr::write_rds(adasyn_w_triggers_train,"./data/interim/adasyn_w_triggers_train.rds", compress = "bz2", compression = 9)
adasyn_w_triggers_train$fraud = as.factor(adasyn_w_triggers_train$fraud)
# Display class distribution after balancing
print(adasyn_w_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```

```{r}
# Apply ADASYN to dataset WITH trigger features for test set
newData <- ADAS(X = data_w_triggers_test,data_w_triggers_test$fraud,K = 5)

adasyn_w_triggers_test <- newData$data %>% dplyr::select(-c("class"))

# Save balanced test set with triggers
readr::write_rds(adasyn_w_triggers_test,"./data/interim/adasyn_w_triggers_test.rds", compress = "bz2", compression = 9)
adasyn_w_triggers_test$fraud = as.factor(adasyn_w_triggers_test$fraud)
# Display class distribution after balancing
print(adasyn_w_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### SMOTE

```{r}
# Calculate imbalance ratio for training set with triggers
nrow(data_w_triggers_train)/sum(data_w_triggers_train$fraud)

# Apply SMOTE to dataset WITH trigger features for training set
smote_w_triggers_train <- SMOTE(data_w_triggers_train,data_w_triggers_train$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

readr::write_rds(smote_w_triggers_train,"./data/interim/smote_w_triggers_train.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_w_triggers_train$fraud = as.factor(smote_w_triggers_train$fraud)
# Verify class balance after SMOTE
print(smote_w_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```



```{r}
# Calculate imbalance ratio for test set with triggers
nrow(data_w_triggers_test)/sum(data_w_triggers_test$fraud)

# Apply SMOTE to dataset WITH trigger features for test set
smote_w_triggers_test <- SMOTE(data_w_triggers_test,data_w_triggers_test$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

readr::write_rds(smote_w_triggers_test,"./data/interim/smote_w_triggers_test.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_w_triggers_test$fraud = as.factor(smote_w_triggers_test$fraud)
# Verify class balance after SMOTE
print(smote_w_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```



### MWMOTE

```{r}
# Apply MWMOTE to dataset WITH trigger features for training set
newData <- mwmote(dataset = data_w_triggers_train,numInstances = nrow(data_w_triggers_train)-2*sum(data_w_triggers_train$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

mwmote_w_triggers_train<-rbind(data_w_triggers_train,newData)

readr::write_rds(mwmote_w_triggers_train,"./data/interim/mwmote_w_triggers_train.rds", compress = "bz2", compression = 9)
mwmote_w_triggers_train$fraud = as.factor(mwmote_w_triggers_train$fraud)
# Display class distribution after MWMOTE balancing
print(mwmote_w_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```


```{r}
# Apply MWMOTE to dataset WITH trigger features for test set
newData <- mwmote(dataset = data_w_triggers_test,numInstances = nrow(data_w_triggers_test)-2*sum(data_w_triggers_test$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

mwmote_w_triggers_test<-rbind(data_w_triggers_test,newData)

readr::write_rds(mwmote_w_triggers_test,"./data/interim/mwmote_w_triggers_test.rds", compress = "bz2", compression = 9)
mwmote_w_triggers_test$fraud = as.factor(mwmote_w_triggers_test$fraud)
# Display class distribution after MWMOTE balancing
print(mwmote_w_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### ROSE

```{r}
# Apply ROSE to dataset WITH trigger features for training set
#chose seed 100 for reproducibility
#The N variable is the total number that will be generated
#N should essentially be sum of both fraud and non fraud classes
rose_w_triggers_train <- ovun.sample(fraud~. , data = data_w_triggers_train, method = "over",N=(nrow(data_w_triggers_train)-sum(data_w_triggers_train$fraud))*2,seed = 100)$data

readr::write_rds(rose_w_triggers_train,"./data/interim/rose_w_triggers_train.rds", compress = "bz2", compression = 9)
rose_w_triggers_train$fraud = as.factor(rose_w_triggers_train$fraud)
# Display and visualize class distribution after ROSE balancing
table(rose_w_triggers_train$fraud)
barplot(prop.table(table(rose_w_triggers_train$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```



```{r}
# Apply ROSE to dataset WITH trigger features for test set
#chose seed 100 for reproducibility
#The N variable is the total number that will be generated
#N should essentially be sum of both fraud and non fraud classes
rose_w_triggers_test <- ovun.sample(fraud~. , data = data_w_triggers_test, method = "over",N=(nrow(data_w_triggers_test)-sum(data_w_triggers_test$fraud))*2,seed = 100)$data

readr::write_rds(rose_w_triggers_test,"./data/interim/rose_w_triggers_test.rds", compress = "bz2", compression = 9)
rose_w_triggers_test$fraud = as.factor(rose_w_triggers_test$fraud)
# Display and visualize class distribution after ROSE balancing
table(rose_w_triggers_test$fraud)
barplot(prop.table(table(rose_w_triggers_test$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```


# ML Model fit comparisons

This section compares multiple machine learning models (Decision Tree, Random Forest, XGBoost, GLM, Naive Bayes, GBM) across different data balancing techniques (ADASYN, SMOTE, MWMOTE, ROSE) with and without trigger features. Each model is evaluated using ROC-AUC, PR-AUC, and classification metrics.

## Decision Tree

### ADASYN

```{r}
# Train decision tree classifier on ADASYN-balanced data
tree <- rpart(fraud ~ ., data = adasyn_wo_triggers_train, method = 'class', minsplit = 5)
# Generate probability predictions on test set
tree.pred <- predict(tree, adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "prob")[, 2]
roc_obj <- roc(adasyn_wo_triggers_test$fraud, tree.pred)
# Determine optimal classification threshold from ROC curve
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate performance metrics and store results
c <- confusionMatrix(as.factor(tree.pred), adasyn_wo_triggers_test$fraud, mode = "everything", positive = "1")
c <- rbind(c, c("DT adasyn_wo_triggers_train", c$overall[1], c$byClass[c(1,2,5,6,7)], roc_obj$auc, pr.curve(tree.pred, adasyn_wo_triggers_test$fraud)$auc.integral))
c = c[-1,]
# Visualize decision tree structure
rpart.plot(tree, extra = 101)

```


```{r}
# Train decision tree on ADASYN-balanced data WITH trigger features
tree <- rpart(fraud~., data = adasyn_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,adasyn_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(adasyn_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate and store performance metrics
c1 = confusionMatrix(as.factor(tree.pred),adasyn_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT adasyn_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, adasyn_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)


```



### SMOTE

```{r}
# Train decision tree on SMOTE-balanced data WITHOUT trigger features
tree <- rpart(fraud~., data = smote_wo_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,smote_wo_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(smote_wo_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate and store performance metrics
c1 = confusionMatrix(as.factor(tree.pred),smote_wo_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT smote_wo_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, smote_wo_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


```{r}
# Train decision tree on SMOTE-balanced data WITH trigger features
tree <- rpart(fraud~., data = smote_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,smote_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(smote_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate and store performance metrics
c1 = confusionMatrix(as.factor(tree.pred),smote_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT smote_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, smote_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


### MWMOTE

```{r}
# Train decision tree on MWMOTE-balanced data WITHOUT trigger features
tree <- rpart(fraud~., data = mwmote_wo_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,mwmote_wo_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(mwmote_wo_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate and store performance metrics
c1 = confusionMatrix(as.factor(tree.pred),mwmote_wo_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT mwmote_wo_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, mwmote_wo_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


```{r}
# Train decision tree on MWMOTE-balanced data WITH trigger features
tree <- rpart(fraud~., data = mwmote_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,mwmote_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(mwmote_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate and store performance metrics
c1 = confusionMatrix(as.factor(tree.pred),mwmote_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT mwmote_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, mwmote_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)


```


### ROSE

```{r}
# Train decision tree on ROSE-balanced data WITHOUT trigger features
tree <- rpart(fraud~., data = rose_wo_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,rose_wo_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(rose_wo_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate and store performance metrics
c1 = confusionMatrix(as.factor(tree.pred),rose_wo_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT rose_wo_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, rose_wo_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


```{r}
# Train decision tree on ROSE-balanced data WITH trigger features
tree <- rpart(fraud~., data = rose_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,rose_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(rose_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

# Calculate and store performance metrics
c1 = confusionMatrix(as.factor(tree.pred),rose_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT rose_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, rose_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)


```


## Random Forest


### ADASYN

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = adasyn_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = adasyn_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(adasyn_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = adasyn_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = adasyn_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(adasyn_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)

  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```


### SMOTE

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = smote_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = smote_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(smote_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))
rm(c1)

# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = smote_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = smote_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(smote_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(smote_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```


### MWMOTE

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = mwmote_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = mwmote_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(mwmote_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = mwmote_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = mwmote_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(mwmote_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```


### ROSE

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = rose_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = rose_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(rose_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = rose_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = rose_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(rose_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(rose_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



## XG Boost

### ADASYN

```{r}
library(xgboost) # for xgboost

# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(adasyn_wo_triggers_train %>% dplyr::select(-c(fraud))), label= adasyn_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(adasyn_wo_triggers_test%>% dplyr::select(-c(fraud))), label= adasyn_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(adasyn_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```


```{r}

# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(adasyn_w_triggers_train %>% dplyr::select(-c(fraud))), label= adasyn_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(adasyn_w_triggers_test%>% dplyr::select(-c(fraud))), label= adasyn_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))

rm(c1)

xgb.plot.importance(xgb.importance(names(adasyn_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

### SMOTE

```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(smote_wo_triggers_train %>% dplyr::select(-c(fraud))), label= smote_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(smote_wo_triggers_test%>% dplyr::select(-c(fraud))), label= smote_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(smote_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```


```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(smote_w_triggers_train %>% dplyr::select(-c(fraud))), label= smote_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(smote_w_triggers_test%>% dplyr::select(-c(fraud))), label= smote_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(smote_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

### MWMOTE

```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(mwmote_wo_triggers_train %>% dplyr::select(-c(fraud))), label= mwmote_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(mwmote_wo_triggers_test%>% dplyr::select(-c(fraud))), label= mwmote_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(mwmote_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```


```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(mwmote_w_triggers_train %>% dplyr::select(-c(fraud))), label= mwmote_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(mwmote_w_triggers_test%>% dplyr::select(-c(fraud))), label= mwmote_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(mwmote_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

### ROSE

```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(rose_wo_triggers_train %>% dplyr::select(-c(fraud))), label= rose_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(rose_wo_triggers_test%>% dplyr::select(-c(fraud))), label= rose_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(rose_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))


```


```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(rose_w_triggers_train %>% dplyr::select(-c(fraud))), label= rose_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(rose_w_triggers_test%>% dplyr::select(-c(fraud))), label= rose_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(rose_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

## GLM

### ADASYN


```{r}


# Fit the GLM model
fit <- glm(fraud ~ ., data = adasyn_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = adasyn_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = adasyn_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


### SMOTE

```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = smote_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = smote_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = smote_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = smote_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```



### MWMOTE


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = mwmote_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = mwmote_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = mwmote_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


test_prob = predict(fit, newdata = mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")
test_roc = roc(mwmote_w_triggers_test$fraud ~ test_prob, plot = TRUE, print.auc = TRUE)

```


### ROSE

```{r}


# Fit the GLM model
fit <- glm(fraud ~ ., data = rose_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = rose_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


test_prob = predict(fit, newdata = rose_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

test_roc = roc(rose_wo_triggers_test$fraud ~ test_prob, plot = TRUE, print.auc = TRUE)

```


```{r}


# Fit the GLM model
fit <- glm(fraud ~ ., data = rose_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = rose_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_w_triggers_test$fraud, test_prob,curve = T)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > mean(test_prob), 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

test_prob = predict(fit, newdata = rose_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")
test_roc = roc(rose_w_triggers_test$fraud ~ test_prob, plot = TRUE, print.auc = TRUE)

```


## Naive Bayes Classifier

### ADASYN


```{r}

# fit model
fit <- naiveBayes(fraud~., data=adasyn_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(adasyn_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(adasyn_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(adasyn_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=adasyn_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=adasyn_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(adasyn_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(adasyn_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(adasyn_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


### SMOTE


```{r}

# fit model
fit <- naiveBayes(fraud~., data=smote_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=smote_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(smote_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(smote_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(smote_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=smote_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=smote_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(smote_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(smote_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(smote_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```


### MWMOTE


```{r}

# fit model
fit <- naiveBayes(fraud~., data=mwmote_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=mwmote_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(mwmote_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(mwmote_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(mwmote_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=mwmote_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(mwmote_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(mwmote_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(mwmote_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```




### ROSE

```{r}


# fit model
fit <- naiveBayes(fraud~., data=rose_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=rose_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(rose_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(rose_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(rose_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=rose_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=rose_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(rose_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(rose_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(rose_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```



## GBM

### ADASYN

```{r}
# Fitting model

adasyn_wo_triggers_train$fraud = as.numeric(adasyn_wo_triggers_train$fraud) - 1
adasyn_wo_triggers_test$fraud = as.numeric(adasyn_wo_triggers_test$fraud) - 1

# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = adasyn_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(adasyn_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)
```



```{r}
# Fitting model

adasyn_w_triggers_train$fraud = as.numeric(adasyn_w_triggers_train$fraud) - 1
adasyn_w_triggers_test$fraud = as.numeric(adasyn_w_triggers_test$fraud) - 1

# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = adasyn_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = adasyn_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(adasyn_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```




### SMOTE

```{r}
# Fitting model

smote_wo_triggers_train$fraud = as.numeric(smote_wo_triggers_train$fraud) - 1
smote_wo_triggers_test$fraud = as.numeric(smote_wo_triggers_test$fraud) - 1

# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = smote_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = smote_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(smote_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```



```{r}
# Fitting model

smote_w_triggers_train$fraud = as.numeric(smote_w_triggers_train$fraud) - 1
smote_w_triggers_test$fraud = as.numeric(smote_w_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = smote_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = smote_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(smote_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```




### MWMOTE


```{r}
# Fitting model

mwmote_wo_triggers_train$fraud = as.numeric(mwmote_wo_triggers_train$fraud) - 1
mwmote_wo_triggers_test$fraud = as.numeric(mwmote_wo_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = mwmote_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = mwmote_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(mwmote_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```



```{r}
# Fitting model

mwmote_w_triggers_train$fraud = as.numeric(mwmote_w_triggers_train$fraud) - 1
mwmote_w_triggers_test$fraud = as.numeric(mwmote_w_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = mwmote_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(mwmote_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```




### ROSE

```{r}
# Fitting model

rose_wo_triggers_train$fraud = as.numeric(rose_wo_triggers_train$fraud) - 1
rose_wo_triggers_test$fraud = as.numeric(rose_wo_triggers_test$fraud) - 1



# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = rose_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = rose_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(rose_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```



```{r}
# Fitting model

rose_w_triggers_train$fraud = as.numeric(rose_w_triggers_train$fraud) - 1
rose_w_triggers_test$fraud = as.numeric(rose_w_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = rose_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = rose_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(rose_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

# Format results dataframe for presentation and analysis
c = as.data.frame(c)
colnames(c)[1] = "Method+dataset"
colnames(c)[8] = "ROC-AUC"
colnames(c)[9] = "PR-AUC"
# Calculate F2 score (weighted harmonic mean prioritizing recall over precision)
c$F2 = 5*as.numeric(c$Precision)*as.numeric(c$Recall)/((4*as.numeric(c$Precision))+as.numeric(c$Recall))

# Display comprehensive comparative results in formatted table
kable(c , caption = "Comparitive results of ML Models with and w/o triggers")%>%
  kable_styling(bootstrap_options = "striped", "hover", "condensed", "responsive", full_width = F, position = "center") %>% 
 row_spec(0, bold = T, color = "black", background = "yellow") %>% 
 scroll_box( height = "900px",fixed_thead = TRUE)

# Save results for future reference
write_rds(c,"./results/results.rds")

```

```{r}
# Extract metric values from nested structures (flattening to single values)
sapply(c$`Method+dataset`, function(x){x[1]}) -> c$`Method+dataset`
sapply(c$Accuracy, function(x){x[1]}) -> c$Accuracy
sapply(c$Sensitivity, function(x){x[1]}) -> c$Sensitivity
sapply(c$Specificity, function(x){x[1]}) -> c$Specificity
sapply(c$Precision, function(x){x[1]}) -> c$Precision
sapply(c$Recall, function(x){x[1]}) -> c$Recall
sapply(c$F1, function(x){x[1]}) -> c$F1
sapply(c$`ROC-AUC`,function(x){x[1]}) -> c$`ROC-AUC`
sapply(c$`PR-AUC`,function(x){x[1]}) -> c$`PR-AUC`
sapply(c$F2,function(x){x[1]}) -> c$F2
```

```{r}
# Finalize results: convert to numeric, create trigger indicator, and parse method names
c <- 
  c %>% 
  mutate(
    across(.cols = -`Method+dataset`, .fns = as.numeric),
    trigger = ifelse(stringr::str_detect(`Method+dataset`, "wo"), "N", "Y")
  ) %>%
  separate(`Method+dataset`, c("method", "NA_1"), " ", remove = TRUE) %>% 
  separate(NA_1, c("imbalance", NA, NA))

# Export final comparative results to Excel for analysis and reporting
writexl::write_xlsx(c, "./results/final_output.xlsx")

```

