---
title: "SSSIHL-CAS: Insurance Fraud Classifier for Health Insurance, a tool for fraud detection"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: TRUE
    collapsed: yes
    smooth_scroll: yes
  html_notebook: null
---

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: yellow;
    border-color: black;
    color: black;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F,message = F,prompt = F,comment = F)

```


# Project Summary

> Objective:

* This is a tool for insurance fraud detection and prevention using machine learning models. The tool is intended to augment any organizations current Fraud Detection Solution, giving ability to keep pace with the rapidly changing fraud   environment, while minimizing costs and increasing profitability.

> Utility:

* Red flagging claims for possible fraud and further investigation, thereby channelizing the investigation efforts to the   most important/prioritized suspected cases 
* Reducing losses and increasing profitability for insurers / reinsurers 
* Reducing premiums for customer's insurance cover 
* Providing a competitive advantage for the insurers / reinsurers 


# Cleaning the dataset

## Loading required libraries

```{r}
library(readxl)
library(janitor)
library(dplyr)
library(lubridate)
library(caret)
library(rpart)
library(rpart.plot)
library(xgboost)
library(tidyverse)
library(randomForest)
library(pROC)
library(e1071)
library(mlbench)
library(naivebayes)
library(ggplot2)
library(psych)
library(gbm)
library(caTools)
library(glmnet)
library(bruceR)
library(kableExtra)
library(smotefamily)
library(imbalance)
library(ROSE)
library(PRROC)
library(ROCR)

```


# Preparing Input data without triggers

## Loading data and Reducing dimensionality of character columns

```{r}
clean_data1 = readRDS("../Input_dataset/cleaned_input_data.rds")

## Considering top 20 unique values from certain columns with large unique values 

# medical service provider id 
mspid <- 
clean_data1 %>% 
  group_by(medical_service_provider_id) %>% 
  count() %>% 
  arrange(desc(n))

sum(mspid[1:20,2])/sum(mspid[,2])

# removing features that do not contribute in predicting fraud
data_wo_triggers <-
  clean_data1 %>% 
  mutate(medical_service_provider_id = 
           ifelse(medical_service_provider_id %in% mspid[1:20,1]$medical_service_provider_id,
                  medical_service_provider_id, 
                  "Others" )) %>% 
  dplyr::select(-c(insured_id, primary_diagnosis_code,treatment_start_date,treatment_end_date,policy_commencement_date,policy_termination_date,claim_reported_date,claim_count_pa))



```


## Feature scaling

```{r}
k = c("approved_allowed_amount","no_of_days_stayed","birth_date")
data_wo_triggers[k] = scaler(data_wo_triggers[k])

write_rds(data_wo_triggers,"data_wo_triggers.rds",compress = "bz2",compression = 9)

kable(head(data_wo_triggers) , caption = "Data_wo_triggers")%>%
  kable_styling(bootstrap_options = "striped", "hover", "condensed", "responsive", full_width = F, position = "center") %>% 
 row_spec(0, bold = T, color = "black", background = "yellow") %>% 
 scroll_box( height = "900px",fixed_thead = TRUE)

```



## Performing OneHot Encoding to Non-numeric columns

```{r}

dmy <- dummyVars(" ~ .", data = data_wo_triggers, fullRank = TRUE)
data_wo_triggers <- data.frame(predict(dmy, newdata = data_wo_triggers))

```


## checking correlations of cloumns

```{r}
corr = cor(data_wo_triggers)

corr[upper.tri(corr)] = 0
diag(corr) = 0
corr = as.data.frame(corr)
corr = corr %>% tibble::rownames_to_column("colnames_1")
piv = pivot_longer(corr,cols = -colnames_1,  names_to = "colnames", values_to = "correlation")
piv = 
  piv %>% 
  mutate(correlation_1 = abs(correlation)) %>% 
  arrange(desc(correlation_1)) %>% 
  filter(correlation_1 !=0)
#selecting features to drop that are highly correlated
drop = as.vector(subset(piv, correlation_1 > 0.7 , select = colnames))
data_wo_triggers = data_wo_triggers[,!(names(data_wo_triggers) %in% drop$colnames)]
```


### Lasso regression to remove non-influential features in predicting fraud
```{r}


#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(data.matrix(data_wo_triggers %>% dplyr::select(-c("fraud"))), data_wo_triggers$fraud, alpha=1)


#find optimal lambda value that minimizes test MSE
best_lambda = cv_model$lambda.min

#produce plot of test MSE by lambda value
plot(cv_model)

#find coefficients of best model
best_model<-glmnet(data.matrix(data_wo_triggers %>% dplyr::select(-c("fraud"))), data_wo_triggers$fraud, alpha = 1, lambda = best_lambda)
k = as.matrix(coef(best_model))
k = as.data.frame(k[-1,])
k = k %>% rownames_to_column("feature")
names(k)[names(k) == 'k[-1, ]'] <- 'coeff'
k = k[k$coeff!=0,]
k = k[order(abs(k$coeff),decreasing = T),]
if(nrow(k)>30) k = k[1:30,]
data_wo_triggers = data_wo_triggers[,c(k$feature,"fraud")]
```


## Checking for class data Imbalance

```{r}
summary(data_wo_triggers$fraud)
barplot(prop.table(table(data_wo_triggers$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")


```

There is evidence to suggest the existence class data imbalance

## SPlitting Train and Test set

```{r}

# randomly split data in r
set.seed(123)
TrainingIndex = sample(seq_len(nrow(data_wo_triggers)),size = 0.7*nrow(data_wo_triggers))
data_wo_triggers_train <- data_wo_triggers[TrainingIndex,] # Training Set
data_wo_triggers_test <- data_wo_triggers[-TrainingIndex,] # Test Set
```


## Adjusting data for imbalance

### ADASYN

```{r}
newData <- ADAS(X = data_wo_triggers_train,data_wo_triggers_train$fraud,K = 5)

adasyn_wo_triggers_train <- newData$data %>% dplyr::select(-c("class"))


readr::write_rds(adasyn_wo_triggers_train,"adasyn_wo_triggers_train.rds", compress = "bz2", compression = 9)
adasyn_wo_triggers_train$fraud = as.factor(adasyn_wo_triggers_train$fraud)
print(adasyn_wo_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```



```{r}

newData <- ADAS(X = data_wo_triggers_test,data_wo_triggers_test$fraud,K = 5)

adasyn_wo_triggers_test <- newData$data %>% dplyr::select(-c("class"))


readr::write_rds(adasyn_wo_triggers_test,"adasyn_wo_triggers_test.rds", compress = "bz2", compression = 9)
adasyn_wo_triggers_test$fraud = as.factor(adasyn_wo_triggers_test$fraud)
print(adasyn_wo_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### SMOTE

```{r}
nrow(data_wo_triggers_train)/sum(data_wo_triggers_train$fraud)

smote_wo_triggers_train <- SMOTE(data_wo_triggers_train,data_wo_triggers_train$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

readr::write_rds(smote_wo_triggers_train,"smote_wo_triggers_train.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_wo_triggers_train$fraud = as.factor(smote_wo_triggers_train$fraud)
print(smote_wo_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```



```{r}
nrow(data_wo_triggers_test)/sum(data_wo_triggers_test$fraud)

smote_wo_triggers_test <- SMOTE(data_wo_triggers_test,data_wo_triggers_test$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

readr::write_rds(smote_wo_triggers_test,"smote_wo_triggers_test.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_wo_triggers_test$fraud = as.factor(smote_wo_triggers_test$fraud)
print(smote_wo_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```



### MWMOTE

```{r}

newData <- mwmote(dataset = data_wo_triggers_train,numInstances = nrow(data_wo_triggers_train)-2*sum(data_wo_triggers_train$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

mwmote_wo_triggers_train<-rbind(data_wo_triggers_train,newData)

readr::write_rds(mwmote_wo_triggers_train,"mwmote_wo_triggers_train.rds", compress = "bz2", compression = 9)
mwmote_wo_triggers_train$fraud = as.factor(mwmote_wo_triggers_train$fraud)
print(mwmote_wo_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```


```{r}
newData <- mwmote(dataset = data_wo_triggers_test,numInstances = nrow(data_wo_triggers_test)-2*sum(data_wo_triggers_test$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

mwmote_wo_triggers_test<-rbind(data_wo_triggers_test,newData)

readr::write_rds(mwmote_wo_triggers_test,"mwmote_wo_triggers_test.rds", compress = "bz2", compression = 9)
mwmote_wo_triggers_test$fraud = as.factor(mwmote_wo_triggers_test$fraud)
print(mwmote_wo_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### ROSE

```{r}
#chose seed 100
#The N variable is the total number that will be generated
#N should essentially be sum of both fraud and non fraud
rose_wo_triggers_train <- ovun.sample(fraud~. , data = data_wo_triggers_train, method = "over",N=(nrow(data_wo_triggers_train)-sum(data_wo_triggers_train$fraud))*2,seed = 100)$data

readr::write_rds(rose_wo_triggers_train,"rose_wo_triggers_train.rds", compress = "bz2", compression = 9)
rose_wo_triggers_train$fraud = as.factor(rose_wo_triggers_train$fraud)
table(rose_wo_triggers_train$fraud)
barplot(prop.table(table(rose_wo_triggers_train$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```



```{r}
#chose seed 100
#The N variable is the total number that will be generated
#N should essentially be sum of both fraud and non fraud
rose_wo_triggers_test <- ovun.sample(fraud~. , data = data_wo_triggers_test, method = "over",N=(nrow(data_wo_triggers_test)-sum(data_wo_triggers_test$fraud))*2,seed = 100)$data

readr::write_rds(rose_wo_triggers_test,"rose_wo_triggers_test.rds", compress = "bz2", compression = 9)
rose_wo_triggers_test$fraud = as.factor(rose_wo_triggers_test$fraud)
table(rose_wo_triggers_test$fraud)
barplot(prop.table(table(rose_wo_triggers_test$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```


# Preparing input dataset with triggers


## Loading trigger functions

```{r}
source("triggers for package.R")

```


## claim_amount_trigger_map

```{r}
trigger_data <- read_excel("../files_for_trigger_functions/trigger_table.xlsx")

data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, package_amount_trigger)

data_w_triggers <- claim_amount_trigger_map(claims_file = clean_data1,claim_paid_field = "approved_allowed_amount",triggers_file = data,procedure_code = "primary_procedure_code",procedure_amount_field = "package_amount_trigger")
```


## hosp_days_trigger_map

```{r}
data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, admission_days_trigger)
data_w_triggers <- hosp_days_trigger_map(claims_file = data_w_triggers,no_of_days_stayed = "no_of_days_stayed",triggers_file = data,procedure_code = "primary_procedure_code",admission_days_trigger = "admission_days_trigger")
```


## age_trigger_map

```{r}

data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(lower_age_limit, upper_age_limit, primary_procedure_code)

data$primary_procedure_code <- as.character(data$primary_procedure_code)

data_w_triggers <- age_trigger_map(claims_file = data_w_triggers,age_column_name = "birth_date",triggers_file = data,lower_age_limit = "lower_age_limit", upper_age_limit = "upper_age_limit",procedure_code = "primary_procedure_code")

```

## gender_trigger_map

```{r}
data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, gender_trigger)

data_w_triggers <- gender_trigger_map(claims_file = data_w_triggers,gender_column_name = "gender_code",triggers_file = data,procedure_code = "primary_procedure_code",gender_trigger = "gender_trigger")
```


## claim_count_trigger_map

```{r}

data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, claim_count_trigger)

data_w_triggers <- claim_count_trigger_map(claim_count = data_w_triggers,claim_count_pa = "claim_count_pa",claim_count_trigger_file = data,procedure_code_field = "primary_procedure_code",claim_count_trigger = "claim_count_trigger")
```


## close_prox_trigger_map

```{r}

data <- 
  trigger_data %>% 
  janitor::clean_names() %>% 
  dplyr::select(primary_procedure_code, close_prox_days)

data_w_triggers <-  close_prox_trigger_map(claims_file = data_w_triggers,treatment_start_date = "treatment_start_date",policy_commencement_date = "policy_commencement_date",triggers_file = data,procedure_code = "primary_procedure_code",close_prox_days = "close_prox_days")
```


## treatment_date_validity_trigger_map

```{r}

data_w_triggers <- treatment_date_validity_trigger_map(claims_file = data_w_triggers,treatment_start_date_field = "treatment_start_date",policy_commencement_date_field = "policy_commencement_date",policy_termination_date_field = "policy_termination_date")
```


## claim_reported_delay_trigger_map

```{r}

data_w_triggers <- claim_reported_delay_trigger_map(claims_file = data_w_triggers,treatment_end_date_field = "treatment_end_date",claim_reported_date_field = "claim_reported_date",claim_delay_days = 15)
```


## hospital_empanelled_trigger_map

```{r}
data <- 
  read_excel("../files_for_trigger_functions/empanelled_hospitals_list_masked.xlsx") %>% 
  janitor::clean_names() %>% 
  dplyr::select(hosp_id)

data_w_triggers <- hospital_empanelled_trigger_map(claims_file = data_w_triggers,hospital_id_field = "medical_service_provider_id",empanelled_hospitals_list = data,empanelled_hospital_id = "hosp_id")
```



## hospital_distance_trigger_map

```{r}
data <- 
  read_excel("../files_for_trigger_functions/district_distances_masked.xlsx") %>% 
  tidyr::pivot_longer(cols=-"location_1",
                      names_to = "location_2",
                      values_to = "distance")

data_w_triggers <- hospital_distance_trigger_map(claims_file = data_w_triggers,residence_location_field = "residence_location",hospital_location_field = "hospital_location",hospital_distance_file = data,residence_location_map = "location_1",hospital_location_map = "location_2",hospital_distance_field = "distance", distance_threshold = 100)
```


```{r}
skimr::skim(data_w_triggers)
```


# Preparing Input data with triggers

## Reducing dimensionality of character columns

```{r}
## Considering top 20 unique values from certain columns with large unique values 

# medical service provider id 
mspid <- 
data_w_triggers %>% 
  group_by(medical_service_provider_id) %>% 
  count() %>% 
  arrange(desc(n))

sum(mspid[1:20,2])/sum(mspid[,2])

data_w_triggers <-
  data_w_triggers %>% 
  mutate(medical_service_provider_id = 
           ifelse(medical_service_provider_id %in% mspid[1:20,1]$medical_service_provider_id,
                  medical_service_provider_id, 
                  "Others" )) %>% 
  dplyr::select(-c(insured_id, primary_diagnosis_code,treatment_start_date,treatment_end_date,policy_commencement_date,policy_termination_date,claim_reported_date))


```




## Performing OneHot Encoding to Non-numeric columns

```{r}

dmy <- dummyVars(" ~ .", data = data_w_triggers, fullRank = TRUE)
data_w_triggers <- data.frame(predict(dmy, newdata = data_w_triggers))
```


## Removing features with SD 0 and removing trigger information features
```{r}
drop1 = c("package_amount_trigger","admission_days_trigger","lower_age_limit","upper_age_limit","age_flag","gender_triggerFemale","claim_count_trigger","close_prox_days","hospital_empanelled_flag","treatment_date_validity_flag")
data_w_triggers = data_w_triggers %>% select(-c(drop1))
```


## Feature scaling

```{r}
k = c("approved_allowed_amount","no_of_days_stayed","birth_date","claim_count_pa","claim_duration_days","distance")
data_w_triggers[k] = scaler(data_w_triggers[k])

kable(head(data_w_triggers), caption = "data_w_triggers")%>%
  kable_styling(bootstrap_options = "striped", "hover", "condensed", "responsive", full_width = F, position = "center") %>% 
 row_spec(0, bold = T, color = "black", background = "yellow") %>% 
 scroll_box( height = "900px",fixed_thead = TRUE)


write_rds(data_w_triggers,"data_w_triggers.rds",compress = "bz2",compression = 9)
```


## Checking for class data Imbalance

```{r}
summary(data_w_triggers$fraud)
barplot(prop.table(table(data_w_triggers$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")


```


## Splitting train and test data
```{r}
set.seed(123)
data_w_triggers = data_w_triggers[,c(colnames(data_wo_triggers),"claim_amount_flag","hosp_days_flag","gender_flag","claim_count_flag","close_prox_flag","distance","hospital_distance_flag","claim_duration_days","claim_reported_delay_flag")]

# randomly split data in r
TrainingIndex = sample(seq_len(nrow(data_w_triggers)),size = 0.7*nrow(data_w_triggers))
data_w_triggers_train <- data_w_triggers[TrainingIndex,] # Training Set
data_w_triggers_test <- data_w_triggers[-TrainingIndex,] # Test Set
```



## Adjusting data for imbalance

### ADASYN

```{r}
newData <- ADAS(X = data_w_triggers_train,data_w_triggers_train$fraud,K = 5)

adasyn_w_triggers_train <- newData$data %>% dplyr::select(-c("class"))


readr::write_rds(adasyn_w_triggers_train,"adasyn_w_triggers_train.rds", compress = "bz2", compression = 9)
adasyn_w_triggers_train$fraud = as.factor(adasyn_w_triggers_train$fraud)
print(adasyn_w_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```



```{r}
newData <- ADAS(X = data_w_triggers_test,data_w_triggers_test$fraud,K = 5)

adasyn_w_triggers_test <- newData$data %>% dplyr::select(-c("class"))


readr::write_rds(adasyn_w_triggers_test,"adasyn_w_triggers_test.rds", compress = "bz2", compression = 9)
adasyn_w_triggers_test$fraud = as.factor(adasyn_w_triggers_test$fraud)
print(adasyn_w_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### SMOTE

```{r}
nrow(data_w_triggers_train)/sum(data_w_triggers_train$fraud)

smote_w_triggers_train <- SMOTE(data_w_triggers_train,data_w_triggers_train$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

readr::write_rds(smote_w_triggers_train,"smote_w_triggers_train.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_w_triggers_train$fraud = as.factor(smote_w_triggers_train$fraud)
print(smote_w_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```



```{r}
nrow(data_w_triggers_test)/sum(data_w_triggers_test$fraud)

smote_w_triggers_test <- SMOTE(data_w_triggers_test,data_w_triggers_test$fraud ,K = 5, dup_size = 18)$data %>% dplyr::select(-c("class"))

readr::write_rds(smote_w_triggers_test,"smote_w_triggers_test.rds", compress = "bz2", compression = 9)

#new SMOTE balanced data
smote_w_triggers_test$fraud = as.factor(smote_w_triggers_test$fraud)
print(smote_w_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```



### MWMOTE

```{r}
newData <- mwmote(dataset = data_w_triggers_train,numInstances = nrow(data_w_triggers_train)-2*sum(data_w_triggers_train$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

mwmote_w_triggers_train<-rbind(data_w_triggers_train,newData)

readr::write_rds(mwmote_w_triggers_train,"mwmote_w_triggers_train.rds", compress = "bz2", compression = 9)
mwmote_w_triggers_train$fraud = as.factor(mwmote_w_triggers_train$fraud)
print(mwmote_w_triggers_train %>% group_by(fraud) %>% summarise(count = n()))
```


```{r}
newData <- mwmote(dataset = data_w_triggers_test,numInstances = nrow(data_w_triggers_test)-2*sum(data_w_triggers_test$fraud),classAttr = "fraud",kNoisy = 5,kMajority = 3,kMinority = 1,threshold = 5,cmax = 2,cclustering = 3)

mwmote_w_triggers_test<-rbind(data_w_triggers_test,newData)

readr::write_rds(mwmote_w_triggers_test,"mwmote_w_triggers_test.rds", compress = "bz2", compression = 9)
mwmote_w_triggers_test$fraud = as.factor(mwmote_w_triggers_test$fraud)
print(mwmote_w_triggers_test %>% group_by(fraud) %>% summarise(count = n()))
```


### ROSE

```{r}
#chose seed 100
#The N variable is the total number that will be generated
#N should essentially be sum of both fraud and non fraud
rose_w_triggers_train <- ovun.sample(fraud~. , data = data_w_triggers_train, method = "over",N=(nrow(data_w_triggers_train)-sum(data_w_triggers_train$fraud))*2,seed = 100)$data

readr::write_rds(rose_w_triggers_train,"rose_w_triggers_train.rds", compress = "bz2", compression = 9)
rose_w_triggers_train$fraud = as.factor(rose_w_triggers_train$fraud)
table(rose_w_triggers_train$fraud)
barplot(prop.table(table(rose_w_triggers_train$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```



```{r}
#chose seed 100
#The N variable is the total number that will be generated
#N should essentially be sum of both fraud and non fraud
rose_w_triggers_test <- ovun.sample(fraud~. , data = data_w_triggers_test, method = "over",N=(nrow(data_w_triggers_test)-sum(data_w_triggers_test$fraud))*2,seed = 100)$data

readr::write_rds(rose_w_triggers_test,"rose_w_triggers_test.rds", compress = "bz2", compression = 9)
rose_w_triggers_test$fraud = as.factor(rose_w_triggers_test$fraud)
table(rose_w_triggers_test$fraud)
barplot(prop.table(table(rose_w_triggers_test$fraud)),
        col = rainbow(3),
        ylim = c(0, 1),
        main = "Class Distribution")
```


# ML Model fit comparisons

## Decision Tree

### ADASYN

```{r}
tree <- rpart(fraud ~ ., data = adasyn_wo_triggers_train, method = 'class', minsplit = 5)
tree.pred <- predict(tree, adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "prob")[, 2]
roc_obj <- roc(adasyn_wo_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)

c <- confusionMatrix(as.factor(tree.pred), adasyn_wo_triggers_test$fraud, mode = "everything", positive = "1")
c <- rbind(c, c("DT adasyn_wo_triggers_train", c$overall[1], c$byClass[c(1,2,5,6,7)], roc_obj$auc, pr.curve(tree.pred, adasyn_wo_triggers_test$fraud)$auc.integral))
c = c[-1,]
rpart.plot(tree, extra = 101)

```


```{r}
tree <- rpart(fraud~., data = adasyn_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,adasyn_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(adasyn_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)


c1 = confusionMatrix(as.factor(tree.pred),adasyn_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT adasyn_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, adasyn_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)


```



### SMOTE

```{r}
tree <- rpart(fraud~., data = smote_wo_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,smote_wo_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(smote_wo_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)


c1 = confusionMatrix(as.factor(tree.pred),smote_wo_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT smote_wo_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, smote_wo_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


```{r}
tree <- rpart(fraud~., data = smote_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,smote_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(smote_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)


c1 = confusionMatrix(as.factor(tree.pred),smote_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT smote_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, smote_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


### MWMOTE

```{r}
tree <- rpart(fraud~., data = mwmote_wo_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,mwmote_wo_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(mwmote_wo_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)


c1 = confusionMatrix(as.factor(tree.pred),mwmote_wo_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT mwmote_wo_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, mwmote_wo_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


```{r}
tree <- rpart(fraud~., data = mwmote_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,mwmote_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(mwmote_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)


c1 = confusionMatrix(as.factor(tree.pred),mwmote_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT mwmote_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, mwmote_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)


```


### ROSE

```{r}
tree <- rpart(fraud~., data = rose_wo_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,rose_wo_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(rose_wo_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)


c1 = confusionMatrix(as.factor(tree.pred),rose_wo_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT rose_wo_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, rose_wo_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)

```


```{r}
tree <- rpart(fraud~., data = rose_w_triggers_train, method = 'class',minsplit = 5)
tree.pred = predict(tree,rose_w_triggers_test %>% dplyr::select(-c("fraud") ),type = "prob")[,2]
roc_obj <- roc(rose_w_triggers_test$fraud, tree.pred)
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(tree.pred > threshold, 1, 0)


c1 = confusionMatrix(as.factor(tree.pred),rose_w_triggers_test$fraud,mode = "everything",positive = "1")
c = rbind(c,c("DT rose_w_triggers_train",c1$overall[1],c1$byClass[c(1,2,5,6,7)],roc_obj$auc,pr.curve(tree.pred, rose_w_triggers_test$fraud)$auc.integral))
rm(c1)

rpart.plot(tree, extra = 101)


```


## Random Forest


### ADASYN

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = adasyn_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = adasyn_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(adasyn_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = adasyn_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = adasyn_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(adasyn_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)

  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```


### SMOTE

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = smote_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = smote_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(smote_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))
rm(c1)

# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = smote_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = smote_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(smote_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(smote_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```


### MWMOTE

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = mwmote_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = mwmote_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(mwmote_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = mwmote_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = mwmote_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(mwmote_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```


### ROSE

```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = rose_wo_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = rose_wo_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(rose_wo_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



```{r}
# train the random forest model
classifier_rf <- randomForest(fraud ~ ., data = rose_w_triggers_train, ntree = 100)

# predict probabilities on the test set
y_prob <- predict(classifier_rf, newdata = rose_w_triggers_test, type = "prob")[,2]

# calculate ROC AUC and PR AUC
roc_obj <- roc(rose_w_triggers_test$fraud, y_prob)
pr_obj <- pr.curve(scores.class0 = y_prob, weights.class0 = ifelse(rose_w_triggers_test$fraud == "1", 1, 0))

# determine threshold using ROC curve
threshold <- coords(roc_obj, "best")[[1]]

# classify test set based on threshold
y_pred <- ifelse(y_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(y_pred), mode = "everything", positive = "1")
c <- rbind(c, c("RF rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_obj$auc, pr_obj$auc.integral))

rm(c1)  
  
# Plotting model
plot(classifier_rf)
  
# Importance plot
importance(classifier_rf)
  
# Variable importance plot
varImpPlot(classifier_rf,top= 10)
```



## XG Boost

### ADASYN

```{r}
library(xgboost) # for xgboost

# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(adasyn_wo_triggers_train %>% dplyr::select(-c(fraud))), label= adasyn_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(adasyn_wo_triggers_test%>% dplyr::select(-c(fraud))), label= adasyn_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(adasyn_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```


```{r}

# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(adasyn_w_triggers_train %>% dplyr::select(-c(fraud))), label= adasyn_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(adasyn_w_triggers_test%>% dplyr::select(-c(fraud))), label= adasyn_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))

rm(c1)

xgb.plot.importance(xgb.importance(names(adasyn_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

### SMOTE

```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(smote_wo_triggers_train %>% dplyr::select(-c(fraud))), label= smote_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(smote_wo_triggers_test%>% dplyr::select(-c(fraud))), label= smote_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(smote_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```


```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(smote_w_triggers_train %>% dplyr::select(-c(fraud))), label= smote_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(smote_w_triggers_test%>% dplyr::select(-c(fraud))), label= smote_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(smote_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

### MWMOTE

```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(mwmote_wo_triggers_train %>% dplyr::select(-c(fraud))), label= mwmote_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(mwmote_wo_triggers_test%>% dplyr::select(-c(fraud))), label= mwmote_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(mwmote_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```


```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(mwmote_w_triggers_train %>% dplyr::select(-c(fraud))), label= mwmote_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(mwmote_w_triggers_test%>% dplyr::select(-c(fraud))), label= mwmote_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(mwmote_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

### ROSE

```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(rose_wo_triggers_train %>% dplyr::select(-c(fraud))), label= rose_wo_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(rose_wo_triggers_test%>% dplyr::select(-c(fraud))), label= rose_wo_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_wo_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(rose_wo_triggers_train %>% dplyr::select(-c("fraud"))),model = model))


```


```{r}
# Convert the data to DMatrix format
dtrain <- xgb.DMatrix(data = data.matrix(rose_w_triggers_train %>% dplyr::select(-c(fraud))), label= rose_w_triggers_train$fraud)
dtest <- xgb.DMatrix(data = data.matrix(rose_w_triggers_test%>% dplyr::select(-c(fraud))), label= rose_w_triggers_test$fraud)

# Train the model
model <- xgb.train(data = dtrain, nrounds = 15)

# Use the model to make predictions on test data
pred_y <- predict(model, dtest)

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_w_triggers_test$fraud, pred_y)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_w_triggers_test$fraud, pred_y)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
tree.pred <- ifelse(pred_y > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(tree.pred), mode = "everything", positive = "1")
c <- rbind(c, c("XGB rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

xgb.plot.importance(xgb.importance(names(rose_w_triggers_train %>% dplyr::select(-c("fraud"))),model = model))

```

## GLM

### ADASYN


```{r}


# Fit the GLM model
fit <- glm(fraud ~ ., data = adasyn_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = adasyn_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = adasyn_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


### SMOTE

```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = smote_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = smote_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = smote_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = smote_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```



### MWMOTE


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = mwmote_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = mwmote_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# Fit the GLM model
fit <- glm(fraud ~ ., data = mwmote_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


test_prob = predict(fit, newdata = mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")
test_roc = roc(mwmote_w_triggers_test$fraud ~ test_prob, plot = TRUE, print.auc = TRUE)

```


### ROSE

```{r}


# Fit the GLM model
fit <- glm(fraud ~ ., data = rose_wo_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = rose_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


test_prob = predict(fit, newdata = rose_wo_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

test_roc = roc(rose_wo_triggers_test$fraud ~ test_prob, plot = TRUE, print.auc = TRUE)

```


```{r}


# Fit the GLM model
fit <- glm(fraud ~ ., data = rose_w_triggers_train, family = binomial(link = "logit"))

# Testing Set
test_prob <- predict(fit, newdata = rose_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_w_triggers_test$fraud, test_prob,curve = T)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_glm_test_pred <- ifelse(test_prob > mean(test_prob), 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(model_glm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GLM rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

test_prob = predict(fit, newdata = rose_w_triggers_test %>% dplyr::select(-c("fraud")), type = "response")
test_roc = roc(rose_w_triggers_test$fraud ~ test_prob, plot = TRUE, print.auc = TRUE)

```


## Naive Bayes Classifier

### ADASYN


```{r}

# fit model
fit <- naiveBayes(fraud~., data=adasyn_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(adasyn_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(adasyn_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(adasyn_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(adasyn_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=adasyn_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=adasyn_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(adasyn_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(adasyn_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(adasyn_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(adasyn_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


### SMOTE


```{r}

# fit model
fit <- naiveBayes(fraud~., data=smote_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=smote_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(smote_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(smote_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(smote_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(smote_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=smote_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=smote_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(smote_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(smote_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(smote_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(smote_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```


### MWMOTE


```{r}

# fit model
fit <- naiveBayes(fraud~., data=mwmote_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=mwmote_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(mwmote_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(mwmote_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(mwmote_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(mwmote_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=mwmote_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(mwmote_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(mwmote_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(mwmote_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(mwmote_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```




### ROSE

```{r}


# fit model
fit <- naiveBayes(fraud~., data=rose_wo_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=rose_wo_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(rose_wo_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(rose_wo_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(rose_wo_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(rose_wo_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```


```{r}

# fit model
fit <- naiveBayes(fraud~., data=rose_w_triggers_train)

# summarize the fit
#print(fit)

# make predictions on test set
test_prob <- predict(fit, newdata=rose_w_triggers_test %>% dplyr::select(-c("fraud")), type="raw")[,2]

# calculate ROC AUC
roc_auc <- roc(rose_w_triggers_test$fraud, test_prob)$auc

# calculate PR AUC
pr_auc <- pr.curve(rose_w_triggers_test$fraud, test_prob)$auc.integral

# set threshold based on optimal ROC curve threshold
threshold <- coords(roc(rose_w_triggers_test$fraud, test_prob), "best")[[1]]

# make predictions based on threshold
model_nb_test_pred <- ifelse(test_prob > threshold, 1, 0)

# confusion matrix
c1 <- confusionMatrix(rose_w_triggers_test$fraud, as.factor(model_nb_test_pred), mode="everything", positive="1")
c <- rbind(c, c("NB rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```



## GBM

### ADASYN

```{r}
# Fitting model

adasyn_wo_triggers_train$fraud = as.numeric(adasyn_wo_triggers_train$fraud) - 1
adasyn_wo_triggers_test$fraud = as.numeric(adasyn_wo_triggers_test$fraud) - 1

# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = adasyn_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = adasyn_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(adasyn_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM adasyn_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)
```



```{r}
# Fitting model

adasyn_w_triggers_train$fraud = as.numeric(adasyn_w_triggers_train$fraud) - 1
adasyn_w_triggers_test$fraud = as.numeric(adasyn_w_triggers_test$fraud) - 1

# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = adasyn_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = adasyn_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(adasyn_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(adasyn_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(adasyn_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM adasyn_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

```




### SMOTE

```{r}
# Fitting model

smote_wo_triggers_train$fraud = as.numeric(smote_wo_triggers_train$fraud) - 1
smote_wo_triggers_test$fraud = as.numeric(smote_wo_triggers_test$fraud) - 1

# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = smote_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = smote_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(smote_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM smote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```



```{r}
# Fitting model

smote_w_triggers_train$fraud = as.numeric(smote_w_triggers_train$fraud) - 1
smote_w_triggers_test$fraud = as.numeric(smote_w_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = smote_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = smote_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(smote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(smote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(smote_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM smote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```




### MWMOTE


```{r}
# Fitting model

mwmote_wo_triggers_train$fraud = as.numeric(mwmote_wo_triggers_train$fraud) - 1
mwmote_wo_triggers_test$fraud = as.numeric(mwmote_wo_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = mwmote_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = mwmote_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(mwmote_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM mwmote_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```



```{r}
# Fitting model

mwmote_w_triggers_train$fraud = as.numeric(mwmote_w_triggers_train$fraud) - 1
mwmote_w_triggers_test$fraud = as.numeric(mwmote_w_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = mwmote_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = mwmote_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(mwmote_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(mwmote_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(mwmote_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM mwmote_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```




### ROSE

```{r}
# Fitting model

rose_wo_triggers_train$fraud = as.numeric(rose_wo_triggers_train$fraud) - 1
rose_wo_triggers_test$fraud = as.numeric(rose_wo_triggers_test$fraud) - 1



# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = rose_wo_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = rose_wo_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_wo_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_wo_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(rose_wo_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM rose_wo_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)


```



```{r}
# Fitting model

rose_w_triggers_train$fraud = as.numeric(rose_w_triggers_train$fraud) - 1
rose_w_triggers_test$fraud = as.numeric(rose_w_triggers_test$fraud) - 1


# Fit the GBM model
gbm.model <- gbm(fraud ~ ., data = rose_w_triggers_train, distribution = 'bernoulli')

# Testing Set
test_prob <- predict(gbm.model, newdata = rose_w_triggers_test %>% dplyr::select(-c("fraud")), n.trees = 1000, type = "response")

# Get the ROC curve and calculate the ROC AUC
roc_obj <- roc(rose_w_triggers_test$fraud, test_prob)
roc_auc <- auc(roc_obj)

# Get the PR curve and calculate the PR AUC
pr_obj <- pr.curve(rose_w_triggers_test$fraud, test_prob)
pr_auc <- pr_obj$auc.integral

# Set the threshold based on the optimal ROC curve threshold
threshold <- coords(roc_obj, "best")[[1]]
model_gbm_test_pred <- ifelse(test_prob > threshold, 1, 0)

# Confusion Matrix
c1 <- confusionMatrix(as.factor(rose_w_triggers_test$fraud), as.factor(model_gbm_test_pred), mode = "everything", positive = "1")
c <- rbind(c, c("GBM rose_w_triggers_train", c1$overall[1], c1$byClass[c(1,2,5,6,7)], roc_auc, pr_auc))
rm(c1)

c = as.data.frame(c)
colnames(c)[1] = "Method+dataset"
colnames(c)[8] = "ROC-AUC"
colnames(c)[9] = "PR-AUC"
c$F2 = 5*as.numeric(c$Precision)*as.numeric(c$Recall)/((4*as.numeric(c$Precision))+as.numeric(c$Recall))

kable(c , caption = "Comparitive results of ML Models with and w/o triggers")%>%
  kable_styling(bootstrap_options = "striped", "hover", "condensed", "responsive", full_width = F, position = "center") %>% 
 row_spec(0, bold = T, color = "black", background = "yellow") %>% 
 scroll_box( height = "900px",fixed_thead = TRUE)

write_rds(c,"results.rds")

```

```{r}

sapply(c$`Method+dataset`, function(x){x[1]}) -> c$`Method+dataset`
sapply(c$Accuracy, function(x){x[1]}) -> c$Accuracy
sapply(c$Sensitivity, function(x){x[1]}) -> c$Sensitivity
sapply(c$Specificity, function(x){x[1]}) -> c$Specificity
sapply(c$Precision, function(x){x[1]}) -> c$Precision
sapply(c$Recall, function(x){x[1]}) -> c$Recall
sapply(c$F1, function(x){x[1]}) -> c$F1
sapply(c$`ROC-AUC`,function(x){x[1]}) -> c$`ROC-AUC`
sapply(c$`PR-AUC`,function(x){x[1]}) -> c$`PR-AUC`
sapply(c$F2,function(x){x[1]}) -> c$F2
```

```{r}
c <- 
  c %>% 
  mutate(
    across(.cols = -`Method+dataset`, .fns = as.numeric),
    trigger = ifelse(stringr::str_detect(`Method+dataset`, "wo"), "N", "Y")
  ) %>%
  separate(`Method+dataset`, c("method", "NA_1"), " ", remove = TRUE) %>% 
  separate(NA_1, c("imbalance", NA, NA))


writexl::write_xlsx(c, "final_output.xlsx")

```

